---
title: "STAT 109 Project Spring 2019: Asthma Predictors"
output: pdf_document
---


## Requires the following packages to be installed:
#install.packages('alr3')
#install.packages('R330')
#install.packages("glmnet")
#install.packages('ggmap')
#source("http://people.fas.harvard.edu/~mparzen/stat100/model_select.txt")


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Glossary of Data Frames
data - raw csv contents.
mydata - updates to the raw csv such as additional columns or adjustments.
mydata_sub - A subset of the data the removes unneeded columns
xformed_data - The transformed mydata_sub
train.set - The xformed_data used for training.
test.set - The xformed_data used for testing.


```{r}
#load data 
data=read.csv("https://raw.githubusercontent.com/sabderra/stat109/master/data/500_city_asthma_with_aq_housing_census.csv")

# Save the original data and do all the expansion with mydata
mydata = data

#state dummy variables
ntheast=c('CT', 'ME', 'MA', 'NH', 'RI', 'VT', 'NJ', 'NY', 'PA')
midw=c('ND','MN','SD','NE','KS','IA','MO','IL','WI','MI','IN','OH')
south=c('DE', 'FL','GA', 'MD', 'NC', 'SC', 'VA', 'DC', 'WV', 'AL', 'KY', 
        'MS', 'TN', 'AK', 'LA', 'OK', 'TX')
west=c('AZ', 'CO', 'ID', 'MT', 'NV', 'NM', 'UT', 'WY', 'AL', 'CA', 'HW', 'OR', 'WA')

mydata$Midwest<- 0
mydata$Northeast<- 0
mydata$South<- 0
#mydata$West<- 0

n = nrow(mydata)
for(j in 1:n){ #loop over rows
  if (data[j,]$StateAbbr %in% midw){mydata[j,]$Midwest = 1}
  else if(data[j,]$StateAbbr %in% ntheast){mydata[j,]$Northeast = 1}
  else if(data[j,]$StateAbbr %in% south){mydata[j,]$South = 1}
  #else if(mydata_sub[j,]$StateAbbr %in% west){mydata_sub[j,]$West = 1}
}
```


Add a column that maps the AQI into the levels defined by: https://airnow.gov/index.cfm?action=aqibasics.aqi

```{r}
n = nrow(mydata)
mydata$MaxAQI2<-0
for(j in 1:n){ #loop over rows
  if ( ! is.na(mydata[j,]$MaxAQI)) {
    if(mydata[j,]$MaxAQI > 300){mydata[j,]$MaxAQI2 = 6}
    else if((mydata[j,]$MaxAQI >200 ) && (mydata[j,]$MaxAQI<=300)){mydata[j,]$MaxAQI2 = 5}
    else if((mydata[j,]$MaxAQI >150 ) && (mydata[j,]$MaxAQI<=200)){mydata[j,]$MaxAQI2 = 4}
    else if((mydata[j,]$MaxAQI >100 ) && (mydata[j,]$MaxAQI<=150)){mydata[j,]$MaxAQI2 = 3}
    else if((mydata[j,]$MaxAQI >50 ) && (mydata[j,]$MaxAQI<=100)){mydata[j,]$MaxAQI2 = 2}
    else if((mydata[j,]$MaxAQI >=0 ) && (mydata[j,]$MaxAQI<=50)){mydata[j,]$MaxAQI2 = 1}
  } else {
    mydata[j,]$MaxAQI2 = NA
  }
}

```


```{r}
mydata_sub=mydata[ , -which(names(mydata) %in% c("Year", "StateAbbr", "StateDesc", "CityName",                 "CO2ndMax8hr","NO2Mean1hr","Ozone4thMax8hr","SO2Mean1hr", "PM25WeightedMean24hr","PM10Mean24hr","DayswithAQI","DaysCO","DaysNO2","DaysOzone","DaysSO2","DaysPM25","DaysPM10","X90thPercentileAQI"))]

# Previous the following were also removed.
#"GoodDays", "ModerateDays", "UnhealthyforSensitiveGroupsDays", "UnhealthyDays", "VeryUnhealthyDays","HazardousDays"

cat("Original data dimensions:", dim(mydata), "\n")
cat("After purging unneeded columns:", dim(mydata_sub), "\n")
mydata_sub=na.omit(mydata_sub)
cat("After purging na:", dim(mydata_sub), "\n")
```


This plot shows the difference after NAs (blue) cities are omited. While there is a spattering, there is a concentration in the South East of cities that were omitted.

```{r}
# to install: install.packages('ggmap')
library(ggmap)

ggmap::register_google(key="AIzaSyBfw9xeNd1giAnLbpCTabTrnrFx9noHmak")
map = get_googlemap(location = geocode("united states"), source="google",
                    zoom = 4, scale = 2,
                    maptype ='terrain',
                    color = 'color')


df.ommited = mydata[!(mydata$Lat %in% mydata_sub$Lat),]

mapPoints = ggmap(map) + 
  geom_point(aes(x = Lng, y = Lat, size=AsthmaPct), data=df.ommited, alpha=0.5, color="blue") +
  geom_point(aes(x = Lng, y = Lat, size=AsthmaPct), data=mydata_sub, alpha=0.5, color="darkred")

mapPoints

```


```{r}
cor(mydata_sub[c("PhysicalAct", "Obesity", "Access", "PopulationCount", "Lat", "Lng", "SmokingPct", "MedianAQI", "RenterOccupiedPct", "UnempRatePct", "CashPublicAssistPct", "SNAPPct", "NoHealthInsCovPct", "BelowPovLevelPct")])

```

## Transform the predictor variables

```{r}
library(alr3)
library(R330)
```

```{r}
set.seed(123)

# longitude was made positive because powerTransform won't run on negative values 

fit=lm(AsthmaPct~PopulationCount+Lat+abs(Lng)+SmokingPct+Access+Obesity+PhysicalAct+MaxAQI+MaxAQI2+MedianAQI+RenterOccupiedPct+UnempRatePct+CashPublicAssistPct+SNAPPct+NoHealthInsCovPct+BelowPovLevelPct+Midwest+Northeast+South, data=mydata_sub)

powerTransform(fit)

summary(powerTransform( with( mydata_sub, cbind(PopulationCount,Lat,abs(Lng),SmokingPct,Access,Obesity,PhysicalAct,MaxAQI,MaxAQI2,MedianAQI,RenterOccupiedPct,UnempRatePct,CashPublicAssistPct,SNAPPct,NoHealthInsCovPct,BelowPovLevelPct))))

```


Run powerTransform on all the possible predictor variables. The polution/air quality predictor are sparsely populated. Rather than impute them we will are omiting any NA values. This will reduce the remain rows by approx 50% to 107 rows.


```{r}
summary(fit)
```

```{r}

xformed_data = mydata_sub

# Only those variable transformed are shown here, unchanges are copied from above.
#xformed_data$AsthmaPct=(mydata_sub$AsthmaPct)^0.05
xformed_data$logAsthmaPct=log(mydata_sub$AsthmaPct)

xformed_data$PopulationCount=(mydata_sub$PopulationCount)^(-0.8)
xformed_data$Lat=(mydata_sub$Lat)^2
xformed_data$Lng=(abs(mydata_sub$Lng))
xformed_data$Access=mydata_sub$Access^0.25
xformed_data$PhysicalAct=sqrt(mydata_sub$PhysicalAct)
xformed_data$MaxAQI=mydata_sub$MaxAQI^(-0.33)
  # 1/(mydata_sub$MaxAQI)
  # was also log(mydata_sub$MaxAQI)
xformed_data$MedianAQI=mydata_sub$MedianAQI^(0.63)
  # was log(mydata_sub$MedianAQI)

xformed_data$UnempRatePct=log(mydata_sub$UnempRatePct)
xformed_data$CashPublicAssistPct=log(mydata_sub$CashPublicAssistPct)
xformed_data$SNAPPct=sqrt(mydata_sub$SNAPPct)
xformed_data$NoHealthInsCovPct=sqrt(mydata_sub$NoHealthInsCovPct)
xformed_data$BelowPovLevelPct=sqrt(mydata_sub$BelowPovLevelPct)

# Unchanged
#xformed_data$MaxAQI2=log(mydata_sub$MaxAQI2)
#xformed_data$SmokingPct=log(mydata_sub$SmokingPct)
#xformed_data$Obesity=(mydata_sub$Obesity)
#xformed_data$RenterOccupiedPct=(mydata_sub$RenterOccupiedPct)
#xformed_data$Midwest=mydata_sub$Midwest
#xformed_data$Northeast=mydata_sub$Northeast
#xformed_data$South=mydata_sub$South
#West=mydata_sub$West

names(xformed_data)
```


Split the data into separate training and test sets containing 80% and 20% respectfully of the observations.

```{r}
set.seed(123)
sample<-sample.int(n=nrow(xformed_data),
                   size=floor(.80*nrow(xformed_data)),replace=F)
train.set <- xformed_data[sample,] #train data
test.set <- xformed_data[-sample,] #test data

cat("Training set dim:", dim(train.set),"\n")
cat("Test set dim:", dim(test.set),"\n")

```

Note from here on the train.set is used.

```{r}
library(sandwich)
library(car)
library(lmtest)
```

## Fit the transformed variables

```{r}

fit.transform=lm(logAsthmaPct~PopulationCount+Lat+Lng+SmokingPct+Access+Obesity+PhysicalAct+MedianAQI+MaxAQI+MaxAQI2+RenterOccupiedPct+UnempRatePct+CashPublicAssistPct+SNAPPct+NoHealthInsCovPct+BelowPovLevelPct+Midwest+Northeast+South+GoodDays+UnhealthyforSensitiveGroupsDays+UnhealthyDays+VeryUnhealthyDays+HazardousDays, data=train.set)


#overall tests
print("VIF test")
vif(fit.transform)
print("ncvTest")
ncvTest(fit.transform)
print("summary")
summary(fit.transform)
print("coeftest")
coeftest(fit.transform, vcov. = vcovHC(fit.transform, type = "HC1"))
```

```{r}
#Population Count - Needs to be logged, heteroscedastic
shapiro.test(xformed_data$PopulationCount)
hist(xformed_data$PopulationCount)
par(mfrow=c(2,2))
plot(mydata_sub$AsthmaPct~mydata_sub$PopulationCount)
plot(xformed_data$logAsthmaPct~xformed_data$PopulationCount)
```

```{r}
#Barely not normal - probabily fine, this was not transformed.
shapiro.test(xformed_data$SmokingPct)
hist(xformed_data$SmokingPct)
par(mfrow=c(2,2))
plot(mydata_sub$AsthmaPct~mydata_sub$SmokingPct)
plot(xformed_data$logAsthmaPct~xformed_data$SmokingPct)
```

```{r}
#Slightly skewed but probably fine
shapiro.test(xformed_data$Access)
hist(xformed_data$Access)
par(mfrow=c(2,2))
plot(mydata_sub$AsthmaPct~mydata_sub$Access)
plot(xformed_data$logAsthmaPct~xformed_data$Access)
```


```{r}
#Barely not normal - probably ok - this was not transformed.
shapiro.test(xformed_data$Obesity)
hist(xformed_data$Obesity)
par(mfrow=c(2,2))
plot(mydata_sub$AsthmaPct~mydata_sub$Obesity)
plot(xformed_data$logAsthmaPct~xformed_data$Obesity)

```

```{r}
#Barely not normal - probably ok
shapiro.test(xformed_data$PhysicalAct)
hist(xformed_data$PhysicalAct)
par(mfrow=c(2,2))
plot(mydata_sub$AsthmaPct~mydata_sub$PhysicalAct)
plot(xformed_data$logAsthmaPct~xformed_data$PhysicalAct)

```

```{r}
shapiro.test(xformed_data$MaxAQI)
hist(xformed_data$MaxAQI)
par(mfrow=c(2,2))
plot(mydata_sub$AsthmaPct~mydata_sub$MaxAQI)
plot(xformed_data$logAsthmaPct~xformed_data$MaxAQI)

```

```{r}
#Fine - this is discrete data (not continous)
shapiro.test(xformed_data$MaxAQI2)
hist(xformed_data$MaxAQI2)
par(mfrow=c(2,2))
plot(mydata_sub$AsthmaPct~mydata_sub$MaxAQI2)
plot(xformed_data$logAsthmaPct~xformed_data$MaxAQI2)

```

```{r}
shapiro.test(xformed_data$MedianAQI)
hist(xformed_data$MedianAQI)
par(mfrow=c(2,2))
plot(mydata_sub$AsthmaPct~mydata_sub$MedianAQI)
plot(xformed_data$logAsthmaPct~xformed_data$MedianAQI)

```

```{r}
#Barely Not Normal - probably fine
shapiro.test(xformed_data$RenterOccupiedPct)
hist(xformed_data$RenterOccupiedPct)
par(mfrow=c(2,2))
plot(mydata_sub$AsthmaPct~mydata_sub$RenterOccupiedPct)
plot(xformed_data$logAsthmaPct~xformed_data$RenterOccupiedPct)

```

```{r}
#Fairly not normal - but probaby ok
shapiro.test(xformed_data$UnempRatePct)
hist(xformed_data$UnempRatePct)
par(mfrow=c(2,2))
plot(mydata_sub$AsthmaPct~mydata_sub$UnempRatePct)
plot(xformed_data$logAsthmaPct~xformed_data$UnempRatePct)

```

```{r}
#barely not normal, some evidence of heteroscedasicty, but probably ok
shapiro.test(xformed_data$CashPublicAssistPct)
hist(xformed_data$CashPublicAssistPct)
par(mfrow=c(2,2))
plot(mydata_sub$AsthmaPct~mydata_sub$CashPublicAssistPct)
plot(xformed_data$logAsthmaPct~xformed_data$CashPublicAssistPct)

```

```{r}
#barely not normal, maybe a little heteroscedastic, but probably ok
shapiro.test(xformed_data$SNAPPct)
hist(xformed_data$SNAPPct)
par(mfrow=c(2,2))
plot(mydata_sub$AsthmaPct~mydata_sub$SNAPPct)
plot(xformed_data$logAsthmaPct~xformed_data$SNAPPct)

```

```{r}
#fairly not normal and heteroscedastic, so it should be logged, maybe it helped a little
shapiro.test(xformed_data$NoHealthInsCovPct)
hist(xformed_data$NoHealthInsCovPct)
par(mfrow=c(2,2))
plot(mydata_sub$AsthmaPct~mydata_sub$NoHealthInsCovPct)
plot(xformed_data$logAsthmaPct~xformed_data$NoHealthInsCovPct)
```

```{r}
#not normal and heteroscedastic, so probably should be logged but doesn't appear to solve much
shapiro.test(xformed_data$BelowPovLevelPct)
hist(xformed_data$BelowPovLevelPct)
par(mfrow=c(2,2))
plot(mydata_sub$AsthmaPct~mydata_sub$BelowPovLevelPct)
plot(xformed_data$logAsthmaPct~xformed_data$BelowPovLevelPct)

```






## Fix any multicollinearity issues.

Removed UnhealthyDays(12.933190 / 12.719325)
```{r}

fit.transform.vif=lm(logAsthmaPct~PopulationCount+Lat+Lng+SmokingPct+Access+Obesity+PhysicalAct+MedianAQI+MaxAQI+MaxAQI2+RenterOccupiedPct+UnempRatePct+CashPublicAssistPct+SNAPPct+NoHealthInsCovPct+BelowPovLevelPct+Midwest+Northeast+South+GoodDays+UnhealthyforSensitiveGroupsDays+VeryUnhealthyDays+HazardousDays, data=train.set)

vif(fit.transform.vif)

```

Next highest is  PhysicalAct=10.959835 / 10.946968
```{r}

fit.transform.vif=lm(logAsthmaPct~PopulationCount+Lat+Lng+SmokingPct+Access+Obesity+MedianAQI+MaxAQI+MaxAQI2+RenterOccupiedPct+UnempRatePct+CashPublicAssistPct+SNAPPct+NoHealthInsCovPct+BelowPovLevelPct+Midwest+Northeast+South+GoodDays+UnhealthyforSensitiveGroupsDays+VeryUnhealthyDays+HazardousDays, data=train.set)

vif(fit.transform.vif)
```
Next highest is  SNAPPct=10.218472 / 10.228399
```{r}

fit.transform.vif=lm(logAsthmaPct~PopulationCount+Lat+Lng+SmokingPct+Access+Obesity+MedianAQI+MaxAQI+MaxAQI2+RenterOccupiedPct+UnempRatePct+CashPublicAssistPct+NoHealthInsCovPct+BelowPovLevelPct+Midwest+Northeast+South+GoodDays+UnhealthyforSensitiveGroupsDays+VeryUnhealthyDays+HazardousDays, data=train.set)

vif(fit.transform.vif)
ncvTest(fit.transform.vif)
```

## Remove insignificant variables 

Rather than manually iterating and removing the highest p-value until only significant variables remained, model.select function is used to do it automatically.

```{r}
library(lmtest)

fit.transform.sig = model.select(fit.transform.vif, verbose=FALSE)

summary(fit.transform.sig)

vif(fit.transform.sig)

par(mfrow=c(2,2))
plot(fit.transform.sig)

par(mfrow=c(1,1))
plot(fit.transform.sig$residuals)

plot(fit.transform.sig, which=4)

ncvTest(fit.transform.sig)
resettest(fit.transform.sig)
boxCox(fit.transform.sig)
coeftest(fit.transform.sig, vcov. = vcovHC(fit.transform.sig, type = "HC1"))
```



## Final Model

This is the final handcrafted model. Other models based on auto selection will be below and all wil be compared. This model is the same as fit.transform.sig. 

```{r}

fit.final.transform=lm(logAsthmaPct ~ PopulationCount + Lat + SmokingPct + 
    Access + Obesity + MaxAQI + UnempRatePct + CashPublicAssistPct + 
    BelowPovLevelPct + Midwest + Northeast + South, data=train.set)

vif(fit.final.transform)
ncvTest(fit.final.transform)
summary(fit.final.transform)

par(mfrow=c(2,2))
plot(fit.final.transform)

par(mfrow=c(1,1))
plot(fit.final.transform$residuals)
plot(fit.final.transform, which=4)

```

From ncvTest ( pvalue > 0.05), we fail to reject $H_0$ which says the model is homoscedastic.
From resettest (pvalue > 0.05), we fail to reject $H_0$ which says no further transformation is needed.

The residual diagnositc plot looks well spreadout around 0.
Normal Q-Q reflects the model is reasonble normal.

Cook's distance reports the following observations:
  20, 338, 326
  

## Model Testing

```{r}
# Data frame to collect the RMSE test results.
# Initialize dataframe for storing the model test results.
df.rmse.results = data.frame(matrix(ncol = 3, nrow = 0))
colnames(df.rmse.results) = c("Model", "RMSE", "Time")

# Utility function to add a test result to the global
# variable df.rmse.results
add_rmse_test_result = function(name, result) {
  rmse = round(result, digits=4)
  cat(name, " RMSE = ", rmse, "\n")

  current_time = Sys.time()
  x = data.frame("Model"=name, "RMSE"=rmse, "Time"=current_time)
  df.rmse.results <<- rbind(df.rmse.results, x)
  
  return (df.rmse.results)
}
```

```{r}
calc_loocv_rmse = function(model) {
  sqrt(mean((resid(model) / (1 - hatvalues(model))) ^ 2))
}
```

```{r}
null.fit = lm( logAsthmaPct ~ 1, data=train.set)
```

```{r}

rmse.fit.final.transform = sqrt(mean((test.set$AsthmaPct - exp(predict(fit.final.transform,newdata=test.set)))^2))

add_rmse_test_result("fit.final.transform", rmse.fit.final.transform)
```


## AIC
```{r}
fit.transform.aic=step(fit.transform,direction="backward",trace=FALSE)
summary(fit.transform.aic)
vif(fit.transform.aic)
ncvTest(fit.transform.aic)

calc_loocv_rmse(fit.transform.aic)

rmse.aic = sqrt(mean((test.set$AsthmaPct-exp(predict(fit.transform.aic,newdata=test.set)))^2))

add_rmse_test_result("AIC step(backward)", rmse.aic)

```

```{r}
#interaction variable for backward aic but with reduced final aic model 
fit.transform.interaction.for.backward.aic=lm(logAsthmaPct~(PopulationCount + Lat + SmokingPct + 
    Access + Obesity + PhysicalAct + MaxAQI + UnempRatePct + 
    CashPublicAssistPct + BelowPovLevelPct + Midwest + Northeast + 
    South)^2, data=train.set)

fit.transform.interaction.bwd.aic=step(fit.transform.interaction.for.backward.aic,
                                       direction="backward",
                                       trace=FALSE)

summary(fit.transform.interaction.bwd.aic)
vif(fit.transform.interaction.bwd.aic)
ncvTest(fit.transform.interaction.bwd.aic)

calc_loocv_rmse(fit.transform.interaction.bwd.aic)

rmse.aic.interaction.bwd = sqrt(mean((test.set$AsthmaPct-exp(predict(fit.transform.interaction.bwd.aic,newdata=test.set)))^2))

add_rmse_test_result("AIC step(backward) with Interaction", rmse.aic.interaction.bwd)
```

```{r}
fit.transform.aic.fwd = step(
  null.fit,
  scope=list(upper=fit.transform),
  direction = "forward",
  trace=FALSE
)

summary(fit.transform.aic.fwd)
vif(fit.transform.aic.fwd)
ncvTest(fit.transform.aic.fwd)

calc_loocv_rmse(fit.transform.aic.fwd)

rmse.aic.fwd = sqrt(mean((test.set$AsthmaPct-exp(predict(fit.transform.aic.fwd,newdata=test.set)))^2))

add_rmse_test_result("AIC step(forward)", rmse.aic.fwd)

```

```{r}
#interaction variable for forward aic using reduced aic forward model 
fit.transform.interaction.for.fwd.aic.reduced=lm(logAsthmaPct~(SNAPPct + Access + SmokingPct + Midwest + 
    South + Obesity + PopulationCount + PhysicalAct + Lng + UnempRatePct + 
    RenterOccupiedPct + MaxAQI + Northeast + CashPublicAssistPct)^2, data=train.set)

fit.transform.aic.fwd.interaction.reduced = step(
  null.fit,
  scope=list(upper=fit.transform.interaction.for.fwd.aic.reduced),
  direction = "forward",
  trace=FALSE
)

summary(fit.transform.aic.fwd.interaction.reduced)
vif(fit.transform.aic.fwd.interaction.reduced)
ncvTest(fit.transform.aic.fwd.interaction.reduced)

calc_loocv_rmse(fit.transform.aic.fwd.interaction.reduced)

rmse.aic.fwd.interaction.reduced = sqrt(mean((test.set$AsthmaPct-exp(predict(fit.transform.aic.fwd.interaction.reduced,newdata=test.set)))^2))

add_rmse_test_result("AIC step(forward) w/ interactions reduced", rmse.aic.fwd.interaction.reduced)
```

```{r}
#interaction variables using forward but full model instead of final reduced forward model 
fit.transform.interaction=lm(logAsthmaPct~(PopulationCount+Lat+Lng+SmokingPct+Access+Obesity+PhysicalAct+MedianAQI+MaxAQI+MaxAQI2+RenterOccupiedPct+UnempRatePct+CashPublicAssistPct+SNAPPct+NoHealthInsCovPct+BelowPovLevelPct+Midwest+Northeast+South+GoodDays+UnhealthyforSensitiveGroupsDays+UnhealthyDays+VeryUnhealthyDays+HazardousDays)^2, data=train.set)
  
fit.transform.aic.fwd.interaction.full = step(
  null.fit,
  scope=list(upper=fit.transform.interaction),
  direction = "forward",
  trace=FALSE
)

summary(fit.transform.aic.fwd.interaction.full)
vif(fit.transform.aic.fwd.interaction.full)
ncvTest(fit.transform.aic.fwd.interaction.full)

calc_loocv_rmse(fit.transform.aic.fwd.interaction.full)

rmse.aic.fwd.interaction.full = sqrt(mean((test.set$AsthmaPct-exp(predict(fit.transform.aic.fwd.interaction.full,newdata=test.set)))^2))

add_rmse_test_result("AIC step(forward) w/ interactions full", rmse.aic.fwd.interaction.full)
```

```{r}
fit.transform.aic.both = step(
  null.fit,
  scope=list(upper=fit.transform), 
  direction = "both",
  trace = FALSE
  )

summary(fit.transform.aic.both)
vif(fit.transform.aic.both)
ncvTest(fit.transform.aic.both)

rmse.aic.both = sqrt(mean((test.set$AsthmaPct-exp(predict(fit.transform.aic.both,newdata=test.set)))^2))

add_rmse_test_result("AIC step(both)", rmse.aic.both)

```

```{r}
#interaction variables with reduced fit.transform based on aic both
fit.transform.interaction.for.aic.both=lm(logAsthmaPct~(SNAPPct + Access + SmokingPct + Midwest + 
    South + Obesity + PopulationCount + PhysicalAct + UnempRatePct + 
    RenterOccupiedPct + MaxAQI + Northeast + CashPublicAssistPct)^2, data=train.set)

fit.transform.interaction.aic.both.reduced = step(
  null.fit,
  scope=list(upper=fit.transform.interaction.for.aic.both), 
  direction = "both",
  trace = FALSE
  )

summary(fit.transform.interaction.aic.both.reduced)
vif(fit.transform.interaction.aic.both.reduced)
ncvTest(fit.transform.interaction.aic.both.reduced)

rmse.aic.both.interaction.reduced = sqrt(mean((test.set$AsthmaPct-exp(predict(fit.transform.interaction.aic.both.reduced,newdata=test.set)))^2))

add_rmse_test_result("AIC step(both) w/ interactions reduced", rmse.aic.both.interaction.reduced)
```

```{r}
#interaction variables with full fit.transform based on aic both
fit.transform.aic.both.interaction.full = step(
  null.fit,
  scope=list(upper=fit.transform.interaction), 
  direction = "both",
  trace = FALSE
  )

summary(fit.transform.aic.both.interaction.full)
vif(fit.transform.aic.both.interaction.full)
ncvTest(fit.transform.aic.both.interaction.full)

rmse.aic.both.interaction.full = sqrt(mean((test.set$AsthmaPct-exp(predict(fit.transform.aic.both.interaction.full,newdata=test.set)))^2))

add_rmse_test_result("AIC step(both) w/ interactions full", rmse.aic.both.interaction.full)
```

## BIC
```{r}
n = length(resid(fit.transform)) 
fit.transform.bic = step(fit.transform, 
                         direction = "backward", 
                         k = log(n), 
                         trace = FALSE)

summary(fit.transform.bic)
vif(fit.transform.bic)
ncvTest(fit.transform.bic)

calc_loocv_rmse(fit.transform.bic)

rmse.bic = sqrt(mean((test.set$AsthmaPct-exp(predict(fit.transform.bic,newdata=test.set)))^2))

add_rmse_test_result("BIC step(backward)", rmse.bic)

```

```{r}
#backward bic with interaction from reduced fit.transform
n = length(resid(fit.transform.interaction)) 

fit.transform.interaction.for.bic.bwd=lm(logAsthmaPct~(Lat + SmokingPct + Obesity + PhysicalAct + 
    MaxAQI + UnempRatePct + BelowPovLevelPct + Midwest + Northeast + 
    South)^2, data=train.set)

fit.transform.interaction.bwd.bic.reduced = step(fit.transform.interaction.for.bic.bwd, 
                         direction = "backward", 
                         k = log(n), 
                         trace = FALSE)

summary(fit.transform.interaction.bwd.bic.reduced)
vif(fit.transform.interaction.bwd.bic.reduced)
ncvTest(fit.transform.interaction.bwd.bic.reduced)

calc_loocv_rmse(fit.transform.interaction.bwd.bic.reduced)

rmse.bic.bwd.interaction.reduced = sqrt(mean((test.set$AsthmaPct-exp(predict(fit.transform.interaction.bwd.bic.reduced,newdata=test.set)))^2))

add_rmse_test_result("BIC step(backward) w/ interaction", rmse.bic.bwd.interaction.reduced)
```

```{r}
n = dim(train.set)[1] 

fit.transform.bic.fwd = step(
  null.fit,
  scope=list(upper=fit.transform),
  direction = "forward",
  k = log(n),
  trace = FALSE
)

summary(fit.transform.bic.fwd)
vif(fit.transform.bic.fwd)
ncvTest(fit.transform.bic.fwd)

calc_loocv_rmse(fit.transform.bic.fwd)

rmse.bic.fwd = sqrt(mean((test.set$AsthmaPct-exp(predict(fit.transform.bic.fwd,newdata=test.set)))^2))

add_rmse_test_result("BIC step(forward)", rmse.bic.fwd)

```

```{r}
#forward bic with interaction variables from reduced fit.transform
n = dim(train.set)[1] 

fit.transform.interaction.for.forward.bic=lm(logAsthmaPct~(SNAPPct + Access + SmokingPct + Midwest + 
    South + Obesity + PopulationCount)^2, data=train.set)

fit.transform.bic.fwd.interaction.reduced = step(
  null.fit,
  scope=list(upper=fit.transform.interaction.for.forward.bic),
  direction = "forward",
  k = log(n),
  trace = FALSE
)

summary(fit.transform.bic.fwd.interaction.reduced)
vif(fit.transform.bic.fwd.interaction.reduced)
ncvTest(fit.transform.bic.fwd.interaction.reduced)

calc_loocv_rmse(fit.transform.bic.fwd.interaction.reduced)

rmse.bic.fwd.interaction.reduced = sqrt(mean((test.set$AsthmaPct-exp(predict(fit.transform.bic.fwd.interaction.reduced,newdata=test.set)))^2))

add_rmse_test_result("BIC step(forward) w/ interaction reduced", rmse.bic.fwd.interaction.reduced)
```

```{r}
#bic transform forward using full fit.transform 
n = dim(train.set)[1] 

fit.transform.bic.fwd.interaction.full = step(
  null.fit,
  scope=list(upper=fit.transform.interaction),
  direction = "forward",
  k = log(n),
  trace = FALSE
)

summary(fit.transform.bic.fwd.interaction.full)
vif(fit.transform.bic.fwd.interaction.full)
ncvTest(fit.transform.bic.fwd.interaction.full)

calc_loocv_rmse(fit.transform.bic.fwd.interaction.full)

rmse.bic.fwd.interaction.full = sqrt(mean((test.set$AsthmaPct-exp(predict(fit.transform.bic.fwd.interaction.full,newdata=test.set)))^2))

add_rmse_test_result("BIC step(forward) w/ interaction full", rmse.bic.fwd.interaction.full)
```


```{r}
n = dim(train.set)[1] 

fit.transform.bic.both = step(
  null.fit,
  scope=list(upper=fit.transform),
  direction = "both",
  k = log(n),
  trace = FALSE
)

summary(fit.transform.bic.both)
vif(fit.transform.bic.both)
ncvTest(fit.transform.bic.both)

calc_loocv_rmse(fit.transform.bic.both)

rmse.bic.both = sqrt(mean((test.set$AsthmaPct-exp(predict(fit.transform.bic.both,newdata=test.set)))^2))

add_rmse_test_result("BIC step(both)", rmse.bic.both)

```

```{r}
#bic both interactions with reduced 

fit.transform.interaction.for.both.bic=lm(logAsthmaPct~(SNAPPct + Access + SmokingPct + Midwest + 
    South + Obesity + PopulationCount)^2, data=train.set)

fit.transform.bic.both.interaction.reduced = step(
  null.fit,
  scope=list(upper=fit.transform.interaction.for.both.bic), 
  direction = "both",
  k = log(n),
  trace = FALSE
  )

summary(fit.transform.bic.both.interaction.reduced)
vif(fit.transform.bic.both.interaction.reduced)
ncvTest(fit.transform.bic.both.interaction.reduced)

rmse.bic.both.interaction.reduced = sqrt(mean((test.set$AsthmaPct-exp(predict(fit.transform.bic.both.interaction.reduced,newdata=test.set)))^2))

add_rmse_test_result("BIC step(both) w/ interaction reduced", rmse.bic.both.interaction.reduced)
```

```{r}
#bic both interactions
fit.transform.bic.both.interaction.full = step(
  null.fit,
  scope=list(upper=fit.transform.interaction), 
  direction = "both",
  k = log(n),
  trace = FALSE
  )

summary(fit.transform.bic.both.interaction.full)
vif(fit.transform.bic.both.interaction.full)
ncvTest(fit.transform.bic.both.interaction.full)

rmse.bic.both.interaction.full = sqrt(mean((test.set$AsthmaPct-exp(predict(fit.transform.bic.both.interaction.full,newdata=test.set)))^2))

add_rmse_test_result("BIC step(both) w/ interaction full", rmse.bic.both.interaction.full)
```

## Ridge Regression

```{r}
library(glmnet)
```

```{r}
x = model.matrix(AsthmaPct~.-logAsthmaPct,train.set)[,-1]
y = train.set$AsthmaPct
cv=cv.glmnet(x,y,alpha=0)
fit.ridge = glmnet(x,y,alpha=0,lambda=cv$lambda.min)
x.test=model.matrix(AsthmaPct~.-logAsthmaPct,test.set)
fits.ridge = x.test%*%coef(fit.ridge)
rmse.ridge = sqrt(mean((test.set$AsthmaPct-fits.ridge)^2))

add_rmse_test_result("Ridge Regression", rmse.ridge)

coef(fit.ridge)

```

```{r}
x = model.matrix(AsthmaPct~(PopulationCount+Lat+Lng+SmokingPct+Access+Obesity+PhysicalAct+MedianAQI+MaxAQI+MaxAQI2+RenterOccupiedPct+UnempRatePct+CashPublicAssistPct+SNAPPct+NoHealthInsCovPct+BelowPovLevelPct+Midwest+Northeast+South+GoodDays+UnhealthyforSensitiveGroupsDays+UnhealthyDays+VeryUnhealthyDays+HazardousDays)^2,train.set)[,-1]
y = train.set$AsthmaPct
cv=cv.glmnet(x,y,alpha=0)
fit.ridge.interaction = glmnet(x,y,alpha=0,lambda=cv$lambda.min)
x.test.interaction=model.matrix(AsthmaPct~(PopulationCount+Lat+Lng+SmokingPct+Access+Obesity+PhysicalAct+MedianAQI+MaxAQI+MaxAQI2+RenterOccupiedPct+UnempRatePct+CashPublicAssistPct+SNAPPct+NoHealthInsCovPct+BelowPovLevelPct+Midwest+Northeast+South+GoodDays+UnhealthyforSensitiveGroupsDays+UnhealthyDays+VeryUnhealthyDays+HazardousDays)^2,test.set)
fits.ridge.interaction = x.test.interaction%*%coef(fit.ridge.interaction)
rmse.ridge.interaction = sqrt(mean((test.set$AsthmaPct-fits.ridge.interaction)^2))

add_rmse_test_result("Ridge Regression with Interaction", rmse.ridge.interaction)

coef(fit.ridge.interaction)
```

## Lasso Regression
```{r}
x = model.matrix(AsthmaPct~.-logAsthmaPct,train.set)[,-1]
y = train.set$AsthmaPct
cv=cv.glmnet(x,y,alpha=1)
fit.lasso = glmnet(x,y,alpha=1,lambda=cv$lambda.min)

coef(fit.lasso)

x.test=model.matrix(AsthmaPct~.-logAsthmaPct,test.set)
fits.lasso = x.test%*%coef(fit.lasso)
rmse.lasso = sqrt(mean((fits.lasso-test.set$AsthmaPct)^2))

add_rmse_test_result("Lasso Regression", rmse.lasso)

```

```{r}
x = model.matrix(AsthmaPct~(PopulationCount+Lat+Lng+SmokingPct+Access+Obesity+PhysicalAct+MedianAQI+MaxAQI+MaxAQI2+RenterOccupiedPct+UnempRatePct+CashPublicAssistPct+SNAPPct+NoHealthInsCovPct+BelowPovLevelPct+Midwest+Northeast+South+GoodDays+UnhealthyforSensitiveGroupsDays+UnhealthyDays+VeryUnhealthyDays+HazardousDays)^2,train.set)[,-1]
y = train.set$AsthmaPct
cv=cv.glmnet(x,y,alpha=1)
fit.lasso.interaction = glmnet(x,y,alpha=1,lambda=cv$lambda.min)

coef(fit.lasso.interaction)

x.test=model.matrix(AsthmaPct~(PopulationCount+Lat+Lng+SmokingPct+Access+Obesity+PhysicalAct+MedianAQI+MaxAQI+MaxAQI2+RenterOccupiedPct+UnempRatePct+CashPublicAssistPct+SNAPPct+NoHealthInsCovPct+BelowPovLevelPct+Midwest+Northeast+South+GoodDays+UnhealthyforSensitiveGroupsDays+UnhealthyDays+VeryUnhealthyDays+HazardousDays)^2,test.set)
fits.lasso.interaction = x.test%*%coef(fit.lasso.interaction)
rmse.lasso.interaction = sqrt(mean((fits.lasso.interaction-test.set$AsthmaPct)^2))

add_rmse_test_result("Lasso Regression with Interaction", rmse.lasso.interaction)
```

## Models results
```{r fig.height = 8, fig.width = 11}

# Sort in descending order by RMSE
df.rmse.results = df.rmse.results[with(df.rmse.results, order(-RMSE)), ]

op <- par(mar=c(5,25,5,5))
bp = barplot(height=df.rmse.results$RMSE, 
        names=df.rmse.results$Model, horiz=TRUE,
        xlab="RMSE",
        main="Model Comparison (RMSE)",
        las=1,
        col="skyblue"
        )
rm(op)

# Add the RMSE as labels for readbility.
text(y=bp, x=df.rmse.results$RMSE, pos=2, labels = df.rmse.results$RMSE)

df.rmse.results

# Save results to CSV
write.csv(df.rmse.results, 
          file = "AsthmaPredictors_RMSE_Results.csv",
          row.names = FALSE)
```
